---
title: "Ch 2. The R Language"
output: html_notebook
---

**Assignment** uses the assignment operator "<-" to create a named object that comprises of other objects. 

**c()** denotes a vector

```{r}
x <- c(2,4,6,8)
x
```
Install some add-on packages that we will need

```{r}
install.packages(c("lavaan","semPlot","corrplot","multcomp"))
```
This data set contains observations from sales and product satisfaction survey.

* It has 500 consumers' answers
* **iProdSAT:** satisfaction with a product
* **iSalesSAT:** satisfaction with sales experience
* **iProdREC:** likelihood to recommend the product
* **iSalesREC:** likelihood to recommend the salesperson
* **Segement:** numerically-coded segment

The function **`factor`** is used to encode a vector as a factor/category. For this data set, we set `Segment` to be a categorical factor variable. Observe Segment is now a factor data type:

```{r}
satData <- read.csv("http://goo.gl/UDv12g")
satData$Segment <- factor(satData$Segment)
head(satData)
```

Next we can plot the corr matrix excluding the categorical `Segment` variable in column 3 by specifying -3 in our slice of `satData`.

`corrplot.mixed(corr)`: Using mixed methods to visualize a correlation matrix.

`cor(x,y = NULL)`: computes the correlation of x and y if these are vectors.
```{r}
library(corrplot)
corrplot.mixed(cor(satData[,-3]))
```
