---
title: "ch-5-comparing-groups"
author: "Sonya Hua"
date: "September 7, 2017"
output: rmarkdown::github_document
---

## Comparing Groups: Tables and Visualizations

```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(echo = TRUE)

# set global display parameters
knitr::opts_chunk$set(fig.width=8, fig.height=5, echo = TRUE, fig.align="center") 
```


One might be interested in group data by segment, region, time period, test groups, etc... In all such cases, we're comparing one group of data to another in order to identify an effect. 

### Data Simulation (Consumer Segmentation Data)

We simulate a data set representative of a consumer segmentation project. The scenario is a subscription-based service (such as cable TV or mag subsription), and we collect data from N= 300 respondents on *age, gender, income, # of children, whether they rent (Y,N), subscribe (Y,N). 

Each respondent has been assigned to 1:4 customer segments: "Suburb Mix", "Urban Hip", "Travelers", or "Moving Up"

Segmentation data is moderately complex and we separate our code into 3 steps:

1) Define the data structure: demographic vars, segment names, and sizes

2) Parametrize the distributions of the demographic vars (i.e. mean, variance, sd, probabilities)

3) Code that iterates over the segments and vars to draw random values depending on the definitions and parameters

#### Define the data structure

First step is to define the variable names, data types, segment names, and sample size of each seg
```{r}
# Variables of segments
segVars <- c("age","gender","income","kids","ownHome", "subscribe")

# Distributions of each variable
segVarType <- c("norm","binom","norm","pois","binom","binom")

# Segment names
segNames <- c("Suburb Mix", "Urban Hip", "Travelers", "Moving Up")

# Size or # of obs per segment
segSize <- c(100,50,80,70)
```

Distributions:

* normal (continuous variables)
* binomial (yes/no variables)
* poisson (count variables)

To help us generate the values per segment and variable, we'll create a *look up tables* of distribution parameters.  By storing values this way, we can easily make quick modifications for future purposes without digging through the code. It's best practice to separate data definitions vs. procedural code by declaring the appropriate variables and reusing it in procedural code. 

We're going to create 2 lookup tables: 1 for means (including mean probabilities), and another for standard deviations for normally distributed vars. Since there's 4 segments and 6 variables, we create 4x6 matrices per table. Each row will represent Seg1, Seg2, etc... Each column will represent a different var. In the case of binom,  we specify the mean probabilities. In poiss vars, since we're using a random poisson distribution, we specify the mean number of kids. 
```{r}
segMeans <- matrix(c(
  40, 0.5, 55000, 2, 0.5, 0.1,
  24, 0.7, 21000, 1, 0.2, 0.2,
  58, 0.5, 64000, 0, 0.7, 0.05,
  36, 0.3, 52000, 2, 0.3, 0.2 ), 
  ncol=length(segVars), byrow=T, dimnames= list(segNames, segVars))  
# dimnames: takes on a list of length 2 with row and column names
# byrow= T: filled by rows, not columns

segMeans
```

```{r}
segSDs <- matrix( c(
  5, NA, 12000, NA, NA, NA,
  2, NA, 5000, NA, NA, NA,
  8, NA, 21000, NA, NA, NA,
  4, NA, 10000, NA, NA, NA),
   ncol=length(segVars), byrow=T, dimnames= list(segNames, segVars))

segSDs
```

We have now fully defined the distribution parameters of each segment. 

#### Iterate over each segment and var to fill in values

Let's plan out our code in pseudocode (sentences organized like code). This helps us conceptualize the requirements of our code and we can easily debug concepts before actual coding:

```
Set up data frame "seg.df" and set a seed for replicating random numbers
For each SEGMENT i in "segNames" {
  Set up a temporary df "this.seg" for this SEGMENT's data
  For each var j in "segVars" {
    Use nested if() conditions on "segVarType[j]" to determine data type for VARIABLE
    If it's a certain distribution, use segMeans[i,j and segSDs[i,j] to draw random data 
    for VARIABLE (within SEGMENT) "segSize[i]" times
  }
  Add this SEGMENT's data ("this.seg") to the first data frame ("seg.df")
}
  }
}

```
```{r}
# Translating the pseudocode into R:

seg.df <- NULL
set.seed(02554)

# Iterate over each segment
for (i in seq_along(segNames)) {
  cat(i, segNames[i], "\n")
  this.seg <- data.frame(matrix(NA, nrow=segSize[i], ncol=length(segVars)))
  
  for (j in seq_along(segVars)) {
    if (segVarType[j] == "norm") {
      this.seg[,j] = rnorm(segSize[i], mean=segMeans[i,j], sd=segSDs[i,j])
    } else if (segVarType[j] =="pois") {
      this.seg[,j] = rpois(segSize[i], lambda=segMeans[i,j])
    } else if (segVarType[j]=="binom") {
      this.seg[,j] = rbinom(segSize[i],size=1, prob=segMeans[i,j])  # size = 1 for 1 trial per value
    } else {
      stop("Bad segment data type: ", segVarType[j])
    }
  }
  seg.df <- rbind(seg.df, this.seg) # combines R objects by Rows
}
```

```{r}
# Add in the labels, names, and types
(names(seg.df) <- segVars)
head(seg.df)
```

```{r out.width=12}

# Factorize the appropriate variables
seg.df$Segment <- factor(rep(segNames, times=segSize)) # Within each elmeent of segNames, replicate it SegSize times
seg.df$ownHome <- factor(seg.df$ownHome, labels= c("ownNO", "ownYes")) # ownNo = 0, ownYes = 1
seg.df$gender <- factor(seg.df$gender, labels=c("Female", "Male"))   # Female = 0, Male = 1
seg.df$subscribe <- factor(seg.df$subscribe, labels = c("subNo", "subYes")) # subNo = 0, subYes = 1

# Check Data
head(seg.df)
library(psych)
library(car)
some(seg.df)
tail(seg.df)

```

```{r}
summary(seg.df)
str(seg.df)
library(psych)
describe(seg.df)
```

As a final step, let's save the data frame as a backup so we can use it again in later chapters. Use "/" for path in windows:
```{r}
setwd("c:/users/sonya/documents/git/r-for-marketing-research-and-analytics")
save(seg.df, file="segdf-Rintro-Ch5.RData")
```


### 4.6 Finding Descriptives By Group

With insights about segments, we can engage or tailor offers in different ways in order to reach them. 

#####How does income and gender vary across segments?

An ad hoc way to do this is with df indexing. For example, to find the mean income for "Moving Up":
```{r}
# mean income for Moving Up segment
mean(seg.df$income[seg.df$Segment == "Moving Up"])

# mean income for Moving Up segment who do not subscribe
mean(seg.df$income[seg.df$Segment == "Moving Up" & seg.df$subscribe == "subNo"])

```
This is tedious to do all the time when we want to evaluate a function by conditions. A more general way to do this is with `by(data, INDICES, FUN)`, which divides `data` into groups for each of the unique values in `INDICES` and then apply the `FUN` function to each group 
```{r}

# mean income by segment
by(seg.df$income, seg.df$Segment, mean)

# mean income by segment and subscription status
by(seg.df$income, list(seg.df$Segment, seg.df$subscribe), mean)
```

However, the output is stringy and not user-friendly. Use `aggregate()` instead which works almost the same as `by` in its list form except we need to include `list()` even if we are grouping by 1 variable. The result is a dataframe where we can save results to an object, which we can then index, write to a file, manipulate in other ways. 
```{r}
aggregate(seg.df$income, list(seg.df$Segment), mean)
aggregate(seg.df$income, list(seg.df$Segment, seg.df$subscribe), mean)

seg.income.mean <- aggregate(seg.df$income, list(seg.df$Segment), mean)
head(seg.df$Segment)
seg.df$segIncome <- seg.income.mean[seg.df$Segment, 2]  #  Indexed mean by segment to look up the appropriate value per row of seg.df

# since seg.df$Segment includes all of the segment rows, we are indexing seg.income.mean by its mean so the return output is seg.df$segment with its mean
```

```{r}
head(seg.income.mean)
```

```{r}
head(seg.df$segIncome)
tail(seg.df$segIncome)
```

When we check the `seg.df` data frame, we see that each row of obs matches its segment mean
```{r}
some(seg.df)
```

Caution: This is bad practice since it can lead to erroneous computations. Let's remove that column by setting its value to `NULL`
```{r}
seg.df$segIncome <- NULL
str(seg.df)
```
#### 5.2.1 Formula Syntax

*Formula specification* provides a standard way to desribe relationships among vars. A formula uses the tilde ~ to separate *response vars* from *explanatory vars* y~x. It can also be interpreted as "y in response to x" or "y is modeled by x" depending on the context
```{r}
aggregate(income ~ Segment, data=seg.df, mean)
```


#### 5.2.2 Descriptives for Two-Way Groups 

A common task in marketing is cross-tabulating, separating customers into groups according to 2+ factors. Formula syntax makes it easy to compute a cross tab just by specifying multiple explanatory variables in the form `y~x1+x2....`. For example:

```{r}
aggregate(income~ Segment + ownHome, data=seg.df, mean)
```

We now have a separate group for each combination of `Segment` and `ownHome`.

A formula can be extended to include as many grouping vars as needed. This data frame can also be saved for reuse:
```{r}
agg.data <- aggregate(income~Segment+income+ownHome+subscribe, data=seg.df, mean)
agg.data[2,]

```

##### How frequent do people own homes per segment?

This calls for combinations of `Segment` and `ownHome` with frequency values. We can compute frequencies using `table(factor1, factor2,...)` to obtain 1-way or 2-way count tables

`with(data, expr, ....)` Evaluate an R *expression* in an environment constructed from *data*, possibly modifying (a copy of) the original data.
```{r}
with(seg.df, table(Segment, ownHome))
```
*Observe*: There are 10 observed customers in the "Urban Hip" segment who own homes and 60 in the "Travelers" segment

##### What is the number of kids in each household by segment?
```{r}
with(seg.df, table(kids, Segment))
```
*Observe*: 80 "Travelers" have no kids. Very few households have more than 4 kids. 

`xtabs(formula = y~x, data=)` creates a contingency table (optionally a sparse matrix) from cross-classifying factors, usually contained in a data frame, using a formula interface. The y variable must be a *numeric variable* and not a factor

We can add together the counts of kids to get the total number of children per segment
```{r}
xtabs(kids~Segment, data=seg.df)
xtabs(income~Segment, data=seg.df)
```

#### 5.2.3 Visualizations by Group: Frequencies and Proportions 

The `lattice` package provides a way of plotting cross tabs with frequencies and proportions. `histogram(formula, data, type)` is similar to `hists()` but understands formula notation including conditioning on factors by separating the plot into multiple panes based on that factor. Conditioning is indicated by the symbol "|". `histogram()` automatically assumes by default percent of total counts (proportions) within each group as our response variable unless a response var is specified in the formula notation. 

```{r}
library(lattice)
histogram(~subscribe | Segment, data=seg.df, layout=c(2,2))
```
*Observe* 

* The "Suburban Mix" segment is least likely to subscribe. It suggests that the company might investigate or improve the product to make it more appealing to this group or else stop marketing to them. 

If we want actual counts rather than proportions, we could include the argument `type="count"`.
```{r}
histogram(~subscribe | Segment, data=seg.df, type="count",
           layout=c(4,1), col=c("burlywood","darkolivegreen"))  # Layout includes 4 columns, 1 row

# layout is a numeric vector of length 2 or 3 giving the number of columns, rows, and pages (optional) in a multipanel display. By default, the number of columns is the number of levels of the first conditioning variable and the number of rows is the number of levels of the second conditioning variable.
```

To condition on 1+ factor, include the factor variable in the conditioning part of the formula with "+"
```{r fig.width=6}
histogram(~subscribe | Segment + ownHome, data=seg.df)
```
We can also plot just "yes" proportions vs. both "yes" and "no" bars. We can wrap `prop.table(data, margin=...)` around the regular `table()` to give us the proportions for each cell with proportion to 1) the entire table (default) 2) row (`margin=1`) or 3) column `('margin=2`). 

##### What is the proportion of subscribers within each segment?

Create a proportion table where `margin=2`. 
```{r}
(subSegment.prop <- prop.table(table(seg.df$subscribe, seg.df$Segment), margin=2)) # subscribe =x-axis, Segment = y-axis
```
To plot just the "yes" values, we can use `barchart()` and selct only the 2nd row of the `prop.table()`:

```{r}
barchart(subSegment.prop[2, ],
         xlab="Subscriber proportion by Segment",
         col="darkolivegreen")
```
*Observe* This chart strongly communicates that the Suburb Mix segment has a very low subscription rate within each segment. 

Note the X-axis starts at 0.5, not at 0, so this can be misleading. We'll adjust that using the `xlim=c(LOW,HIGH)` argument where we specify LOW=0. As best practice, always set your xlim and ylim to 0 as a starting value.
```{r}
barchart(subSegment.prop[2, ],
         xlab="Subscriber proportion by Segment",
         col="darkolivegreen",
         xlim=c(0,.25))
```

There's several ways to plot continous data. For ex, how would we plot income ~ segment in our data? A simple way would be to use `aggregate()` to find the mean income, and then use `barchar()` from lattice package to plot the values:

```{r}
seg.mean <- aggregate(income ~ Segment, data=seg.df, mean)
barchart(income~Segment, seg.mean, col="grey",
         main="Avg. income by Segment")
```

To split this out further by home ownership, 1) we have to aggregate the data to include `Segment + ownHome`. 2) tell `barchart()` to use `ownHome` as a `groups=` grouping variable. 
```{r}


seg.income.agg <- aggregate(income ~ Segment + ownHome, data=seg.df, mean)
barchart(income ~ Segment, 
         groups=ownHome, 
         data=seg.income.agg,
         par.settings=simpleTheme(col=terrain.colors(2))) # added simpleTheme option to par.settings to improve the chart colors

#The appearance of a lattice display depends partly on the “theme” active when the display is plotted 

# simpleTheme: Simple interface to generate a list appropriate as a theme, typically used as the par.settings argument in a high level call 
```

There's a variety of color palettes we can use within RColorBrewer package

`brewer.pal(n, name)` Creates nice looking color palettes especially for thematic maps. `n` is the number of different colors in the palette (minimum 3). `name` is the palette name:


Accent	 8
Dark2	 8
Paired	 12
Pastel1	 9
Pastel2	 8
Set1	 9
Set2	 8
Set3	 12
```{r}

```


```{r fig.height=7}
library(RColorBrewer)
display.brewer.all()
```
```{r}
barchart(income ~ Segment, 
         groups=ownHome, 
         data=seg.income.agg,
         par.settings = simpleTheme(col=  brewer.pal(3,"Set3")))   # Using RColorBrewer's Set3 Palette
```
A more informative plot for comparing values of continuous data is a boxplot. A boxplot is better than a barchart because it shows the distribution of values. 

`boxplot()` works with formula syntax to plot a box-and-whiskers plot by factor. 
```{r}
boxplot(income ~ Segment, data=seg.df, yaxt="n", ylab="Income (k)")
ax.seq <- seq(from=0, 120000, 20000)
axis(side=2, # for y-axis
     at=ax.seq,
     labels= paste("$",ax.seq/1000, "k", sep=""),
     las=1) # horizontal direction for text
```
```{r}
# equivalent content but with better graphics using bwplot() from lattice package. 
# CAUTION: bwplot() uses the model formula in the opposite direction x ~ y vs. y~ x

bwplot(Segment ~ income, data=seg.df, horizontal=T, # y is our factor so horizontal == T for for y
       xlab="Income")

# horizontal = :  Determines which of x and y is to be a factor or shingle (y if TRUE, x otherwise). 
# Defaults to FALSE if x is a factor or shingle, TRUE otherwise
```
*Observe* : Travelrs have a much wider distribution of income. Urban Hip has a smaller distriubtion and lowest income 
```{r fig.width=7}
bwplot(Segment ~ income | ownHome, data=seg.df, horizontal = T, xlab="Income $", layout=c(1,2))
```
*Observe* Distribution for travelers is smaller for people who don't own homes. 